---
title: "244-inclass-lab7"
author: "AnnaClaire Marley"
date: "2/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab 7 - More spatial business {.tabset .tabset-fade}

Attach packages: 
```{r, warning = FALSE, message = FALSE}

library(tidyverse)
library(tmap)
library(sf)
library(spatstat)
library(maptools)
library(sp)
library(raster)
# library(rgdal)
library(gstat)
library(plotKML) # for points to raster (they won't have this...just testing)

```

###Part 1. Hawaii raster intro

```{r}

# Read in the raster data

hi_par <- raster("PAR_CLIM_M.tif")
hi_sst <- raster("SST_LTM.tif")
hi_chl <- raster("CHL_LTM.tif")
  
# Base plots
plot(hi_par)
plot(hi_sst)
plot(hi_chl)

```

Reproject to wgs84
```{r}

wgs84 <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +no_defs"

hi_sst_84 <- projectRaster(hi_sst, crs = wgs84, method = "bilinear")

hi_sst_84

# use aggregate to take out some of the resolution and make it run faster
```

raster::aggregate() for resampling

```{r}

sst_rs <- aggregate(hi_sst_84, fact = 10)
plot(sst_rs)

```

Crop a raster:

```{r}

hi_sst_84@extent # total area of the raster extent 
plot(hi_sst_84)

bounds <- as(extent(-156.2, -154.5, 18.7, 20.4), 'SpatialPolygons')

crs(bounds) <- crs(hi_sst_84)

sst_crop <- crop(hi_sst_84, bounds)

plot(sst_crop)

```

Simple raster math:

Nonsensical variable called "tropicality" that is the sum of PAR + SST +2*ChlA and we want to map that variable

Note: for thematic (categorical) raster data, consider using method = "ngm"

```{r}

hi_par_84 <- projectRaster(hi_par, crs = wgs84, method = "bilinear")

hi_chla_84 <- projectRaster(hi_chl, crs = wgs84, method = "bilinear")

```

```{r}

trop <- hi_par_84 + hi_sst_84 +2*hi_chla_84
plot(trop)

```

Now lets try to look at something using tmap:

```{r}

islands <- read_sf(dsn = 'islands', layer = "Island_boundaries") %>% 
  dplyr::select(Island) %>% 
  st_simplify(dTolerance = 10) %>% 
  st_transform(crs = 4326)

plot(islands)

```

```{r}

tmap_mode("plot")

sst_map <- tm_shape(hi_sst_84) +
  tm_raster(title = "Mean Sea Surface Temperature") +
  tm_layout(bg.color = "navyblue",
            legend.position = c("left","bottom"),
            legend.text.color = "white",
            legend.text.size = 0.5) +
  tm_shape(islands) + 
  tm_fill("darkgreen")

tmap_save(sst_map, "AM_hawaii.png", height = 5)

```

conditional rasters and masking

let's say: we have a sensitive species and we're trying to find habitat that it might like. we know they like warm water (average SST >= 25.6 degrees) and solar (PAR) below 54. 

```{r}

hi_sst_84

extent(hi_sst_84) <- extent(hi_par_84)

cr <- raster(nrow = 822,
             ncol = 1229,
             xmn = -160.4365,
             xmx = -154.5373,
             ymn = 18.7309,
             ymx = 22.44634)

sst_new <- resample(hi_sst_84, cr, method = "bilinear")

compareRaster(sst_new, hi_par_84)

plot(sst_new)
plot(hi_par_84)

# make a cropped version just for Kauai
bounds_main <- as(extent(-159.9, -159.2, 21.7, 22.3), 'SpatialPolygons')

crs(bounds_main) <- crs(sst_new)

par_kauai <- crop(hi_par_84, bounds_main)
sst_kauai <- crop(sst_new, bounds_main)

plot(par_kauai)
plot(sst_kauai)

```

Now we only want to isolate the regions where temp >= 25.4, PAR <54

```{r}

par_hab <- par_kauai
par_hab[par_hab >= 54] <- NA
plot(par_hab)

sst_hab <- sst_kauai
sst_hab[sst_hab < 25.4] <- NA

plot(sst_hab)

# where do these overlap? raster::mask

suit_hab <- mask(sst_hab, par_hab)
plot(suit_hab)

```


### Part 2. point pattern analysis

An analysis of red tree voles in humboldt county

```{r}

voles <- read_sf(dsn = 'redtreevoledata', layer = "ds033") %>% 
  dplyr::select(COUNTY) %>% 
  filter(COUNTY == "HUM") %>% 
  st_transform(crs = 4326)

plot(voles)

# get humboldt county outline
humboldt <- read_sf(dsn = 'redtreevoledata', layer = "california_county_shape_file") %>% 
  filter(NAME == "Humboldt") %>% 
  dplyr::select(NAME)

st_crs(humboldt) <- 4326

plot(humboldt)

tm_shape(humboldt) +
  tm_fill() +
  tm_shape(voles) +
  tm_dots(size = 0.2)

ggplot() +
  geom_sf(data = humboldt) +
  geom_sf(data = voles)

ggsave("humvoles.png",
       units = "in",
       width = 4,
       height = 6,
       dpi = 300)


```

WE want to explore point patterns in a few diff ways

- quadrat analysis

- distance based (neighbor analysis, G-function and k-function)

```{r}

voles_sp <- as(voles, 'Spatial')
voles_ppp <- as(voles_sp, "ppp")

humboldt_sp <- as(humboldt, "Spatial")
humboldt_win <- as(humboldt_sp, "owin")

voles_pb <- ppp(voles_ppp$x, voles_ppp$y, window = humboldt_win)

plot(voles_pb)
```

Quadrat test:

```{r}

vole_qt <- quadrat.test(voles_pb, nx = 5, ny = 10)

# testing the null hypothesis of spatial evenness (although you'll hear it called a test for CSR)

plot(voles_pb)
plot(vole_qt, add = TRUE, cex = 0.4)

```


plotting kernel densities for spatial plots
```{r}
point_density <- density(voles_pb, sigma = 0.02)
plot(point_density)

vole_raster <- raster(point_density, crs = wgs84)

tm_shape(vole_raster) +
  tm_raster()

```

Nearest neighbor 

G-function: considers the distance of each observation to its NEAREST neighbor

K-function: considers how close all neighboring observations are to an event (concentric circles)

```{r}

# G-function
r <- seq(0, 0.15, by = 0.005)

gfunction <- envelope(voles_pb, fun = Gest, r = r, nsim = 20)

plot(gfunction$obs ~ gfunction$r, type = "l", col = "black")
lines(gfunction$theo ~ gfunction$r, type = "l", col = "red")

# K/L function: Asking about how close ALL neighbors are to EVERY event in the spatial window 

r2 <- seq(0, 0.5, by = 0.05)

lfunction <- envelope(voles_pb, fun = Lest, r = r2, nsim = 20, global = TRUE)

plot(lfunction$obs ~ lfunction$r, type = "l", col = "blue")
lines(lfunction$theo ~ lfunction$r, type = "l", col = "red")

```

Diggle-Cressie-Loosmore-Ford test of CSR

```{r}

DCLFTest <- dclf.test(voles_pb, nsim = 30)
DCLFTest

```


###Part 3. spatial interpolation by kriging

```{r}

ks_rain <- read_csv("KSRain2.csv")
ks_sf <- st_as_sf(ks_rain, coords = c("LON", "LAT"), crs = 4326)
plot(ks_sf)

ks_counties <- read_sf(dsn = 'KSCounties', layer = "ks_counties_shapefile")

st_crs(ks_counties) <- 4326
plot(ks_counties)

tm_shape(ks_counties) +
  tm_fill() +
  tm_shape(ks_sf) +
  tm_dots("AMT", size = 0.5)

```

```{r}

ks_sp <- as_Spatial(ks_sf)



```

Make a spatial grid to interpolate values over

```{r}

lat <- seq(37,40, length.out = 200)
long <- seq(-94.6, -102, length.out = 200)

grid <- expand.grid(lon = long, lat = lat) #for creating a spatial grid
grid_sf <- st_as_sf(grid, coords = c("lon","lat"), crs = 4326)
grid_sp <- as_Spatial(grid_sf)

```

Then make a variograph and find the variogram model

```{r}

ks_vgm <- variogram(AMT ~ 1, ks_sp)
plot(ks_vgm)

ks_vgm_fit <- fit.variogram(ks_vgm, model = vgm(nugget = 0.2, psill = 0.8, range = 200, model = "Sph"))

plot(ks_vgm, ks_vgm_fit)
```

Now we can do spatial interpolation

```{r}

ks_krige <- krige(AMT ~ 1, ks_sp, grid_sp, model = ks_vgm_fit)

ks_krige_df <- as.data.frame(ks_krige)

ks_krige_2 <- ks_krige_df %>% 
  rename(lon = coords.x1, lat = coords.x2, predicted = var1.pred, err = var1.var)

rain_predicted <- st_as_sf(ks_krige_2, coords = c("lon", "lat"), crs = 4326)

ks <- read_sf(dsn = "states", layer = "cb_2017_us_state_20m") %>% 
  dplyr::select(NAME) %>% 
  filter(NAME == "Kansas") %>% 
  st_transform(crs = 4326)

plot(rain_predicted)

rain_cropped <- st_intersection(rain_predicted, ks)

plot(rain_cropped)

```














