---
title: "244-inclass-lab7"
author: "AnnaClaire Marley"
date: "2/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab 7 - More spatial business {.tabset .tabset-fade}

Attach packages: 
```{r, warning = FALSE, message = FALSE}

library(tidyverse)
library(tmap)
library(sf)
library(spatstat)
library(maptools)
library(sp)
library(raster)
# library(rgdal)
library(gstat)
library(plotKML) # for points to raster (they won't have this...just testing)

```

###Part 1. Hawaii raster intro

```{r}

# Read in the raster data

hi_par <- raster("PAR_CLIM_M.tif")
hi_sst <- raster("SST_LTM.tif")
hi_chl <- raster("CHL_LTM.tif")
  
# Base plots
plot(hi_par)
plot(hi_sst)
plot(hi_chl)

```

Reproject to wgs84
```{r}

wgs84 <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +no_defs"

hi_sst_84 <- projectRaster(hi_sst, crs = wgs84, method = "bilinear")

hi_sst_84

# use aggregate to take out some of the resolution and make it run faster
```

raster::aggregate() for resampling

```{r}

sst_rs <- aggregate(hi_sst_84, fact = 10)
plot(sst_rs)

```

Crop a raster:

```{r}

hi_sst_84@extent # total area of the raster extent 
plot(hi_sst_84)

bounds <- as(extent(-156.2, -154.5, 18.7, 20.4), 'SpatialPolygons')

crs(bounds) <- crs(hi_sst_84)

sst_crop <- crop(hi_sst_84, bounds)

plot(sst_crop)

```

Simple raster math:

Nonsensical variable called "tropicality" that is the sum of PAR + SST +2*ChlA and we want to map that variable

Note: for thematic (categorical) raster data, consider using method = "ngm"

```{r}

hi_par_84 <- projectRaster(hi_par, crs = wgs84, method = "bilinear")

hi_chla_84 <- projectRaster(hi_chl, crs = wgs84, method = "bilinear")

```

```{r}

trop <- hi_par_84 + hi_sst_84 +2*hi_chla_84
plot(trop)

```

Now lets try to look at something using tmap:

```{r}

islands <- read_sf(dsn = 'islands', layer = "Island_boundaries") %>% 
  dplyr::select(Island) %>% 
  st_simplify(dTolerance = 10) %>% 
  st_transform(crs = 4326)

plot(islands)

```

```{r}

tmap_mode("plot")

sst_map <- tm_shape(hi_sst_84) +
  tm_raster(title = "Mean Sea Surface Temperature") +
  tm_layout(bg.color = "navyblue",
            legend.position = c("left","bottom"),
            legend.text.color = "white",
            legend.text.size = 0.5) +
  tm_shape(islands) + 
  tm_fill("darkgreen")

tmap_save(sst_map, "AM_hawaii.png", height = 5)

```

conditional rasters and masking

let's say: we have a sensitive species and we're trying to find habitat that it might like. we know they like warm water (average SST >= 25.6 degrees) and solar (PAR) below 54. 

```{r}

hi_sst_84

extent(hi_sst_84) <- extent(hi_par_84)

cr <- raster(nrow = 822,
             ncol = 1229,
             xmn = -160.4365,
             xmx = -154.5373,
             ymn = 18.7309,
             ymx = 22.44634)

sst_new <- resample(hi_sst_84, cr, method = "bilinear")

compareRaster(sst_new, hi_par_84)

plot(sst_new)
plot(hi_par_84)

# make a cropped version just for Kauai
bounds_main <- as(extent(-159.9, -159.2, 21.7, 22.3), 'SpatialPolygons')

crs(bounds_main) <- crs(sst_new)

par_kauai <- crop(hi_par_84, bounds_main)
sst_kauai <- crop(sst_new, bounds_main)

plot(par_kauai)
plot(sst_kauai)

```

Now we only want to isolate the regions where temp >= 25.4, PAR <54

```{r}

par_hab <- par_kauai
par_hab[par_hab >= 54] <- NA
plot(par_hab)

sst_hab <- sst_kauai
sst_hab[sst_hab < 25.4] <- NA

plot(sst_hab)

# where do these overlap? raster::mask

suit_hab <- mask(sst_hab, par_hab)
plot(suit_hab)

```







